Laplacian envelope

### 2015-09-05 Notes/laplace.tex
hzgdxccsk s

For $s\gtrsim 1.36423$, we set $M=2$ to find
%
\begin{equation}\label{eq:binaryenv}
  \hat{A}(s) = \frac{1}{1+s}.
\end{equation}
%
This does not match up with \eqref{eq:unienvapprox} at the boundary of the two regions, as this is outside the regime of validity of that approximation.
However, it does match up with the full formula, \eqref{eq:unienv}.

@@

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
         {[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + \varepsilon(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - \varepsilon\alpha(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{2 [\varepsilon(\alpha^{\frac{M}{2}}-1)^2
          + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
          { [M-2(1-\varepsilon)]
            [\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression at $M-2$ at $\varepsilon=0$.

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0\text{ or }1$:
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap.
This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.

@@

Nevertheless, the qualitative message of our findings should be more robust to reality:
%
\begin{itemize}
  \item \emph{Short} timescale $\to$ \emph{intermediate} timescales: transition topology goes \\ short and wide $\to$ long and thin.
  \item \emph{Intermediate} timescale $\to$ \emph{long} timescales: transition probabilities go \\ strong and deterministic $\to$ weak and stochastic.
\end{itemize}
%
This is what we expect to find in the different brain regions that store memories for different timescales.

>>>>

For $s\gtrsim 1.36423$, we set $M=2$ to find
%
\begin{equation}\label{eq:binaryenv}
  \hat{A}(s) = \frac{1}{1+s}.
\end{equation}
%
This does not match up with \eqref{eq:unienvapprox} at the boundary of the two regions, as this is outside the regime of validity of that approximation.
However, it does match up with the full formula, \eqref{eq:unienv} at the boundary.

@@

The $\etw_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
         {[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + \varepsilon(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - \varepsilon\alpha(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{2 [\varepsilon(\alpha^{\frac{M}{2}}-1)^2
          + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
          { [M-2(1-\varepsilon)]
            [\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression for $M-2$ at $\varepsilon=0$.

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by asking at what value of $s$ is the optimal value of $\varepsilon=0$ or $1$?
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap.
This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not shorten by one state at a time, but several outgoing transition probabilities are reduced at once.

@@

Nevertheless, the qualitative message of our findings should be more robust to reality:
%
\begin{itemize}
  \item \emph{Short} timescale $\to$ \emph{intermediate} timescales: transition topology goes from \\ short and wide $\to$ long and thin.
  \item \emph{Intermediate} timescale $\to$ \emph{long} timescales: transition probabilities go from \\ strong and deterministic $\to$ weak and stochastic.
\end{itemize}
%
This is what we expect to find in the different brain regions that store memories for different timescales.
